Challenging Fuel Cycle Modeling Assumptions: Facility and Time Step Discretization Effects

### 2015-10-20 paper.tex
add figure captions and refs in text. zoom in on sync-cycle fig

\begin{figure}[h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/sync-cycle.eps}
    \caption[The cycle staggering effect]{
        In this scenario, reactors are deployed annualy with a refueling cycle
        length (including outage) of 18 months.  So all reactors deployed
        every 3rd year refuel together.  All initial 100 LWRs start out with
        their cycles synced as well.  While it may look like separate curves,
        generated power curve (purple) represents only a single time series with many discontinuities.
    }
    \label{fig:sync-cycle}
\end{figure}

@@

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases MF and QF.  However, due to the longer
time steps in case QF, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This increases the discrepency between the two
cases. Eventually the shortage in case MF ends slightly before it does in case
QF. This feedback effect occurs because the reactors have a
conversion ratio greater than 1.0.

The horizontal striations visible in the monthly flows for case MI in Figure
\ref{fig:puinv} are a characteristic of the synchronization of reactor
refueling caused by the shortage.  They begin at the end of the fuel shortage
window when many reactors come online together and are resonances that the
inventory level jumps between.  These resonances gradually disappear as the
simulation progresses farther in time beyond the shortage and refueling cycles
become more staggered.

The in-flows shown for cases MF and QF show a periodic pulsing that begins
during the shortage in about year 120.  This pusling actually has nothing to
do with the shortage and is caused by the retirement of fast reactors.  The
first fast reactors are deployed in year 35 and have an 80 year lifetime.
When they retire, they discharge an entire core's worth of fuel rather than a
single batch.  After a 7-year cooling period, the first discharged full cores
begin to make it through the recycle loop in year 122.  Another interesting
feature visible in all four cases in Figure \ref{fig:puinv} is the minority
fraction of out-flow points that are higher than the bulk majority. These
higher out-flows occur once every 21 months when new reactors are built because
new reactors require a full core of fresh fuel rather than the single batch
needed when just refueling.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-MI.eps}
    \caption[Separated Pu inventory and flow: Case MI Zoom]{
        TODO
    }
    \label{fig:puinv-MI}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-MF.eps}
    \caption[Separated Pu inventory and flow: Case MF Zoom]{
        TODO
    }
    \label{fig:puinv-MF}
\end{figure*}

Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated plutonium
inventory during the shortage generally indicates fewer unfueled reactors.
Lower inventory levels mean available Pu is being utilized more efficiently
rather than sitting idle.  Cases QI and QF with their longer time steps show
higher inventories indicating a more severe shortage.  Post shortage, cases QI and QF
also fall a bit behind with respect to building up Pu stocks due to the delayed
contributions of more unfueled reactors to the separated Pu pool.  The
reactors in all the scenarios request fresh fuel only when they need it. While
this is an unrealistic behavior for real-world facilities, it is closer to a
globally optimum fuel management strategy.  Keeping on-hand fresh fuel batches
would also have the same effect of increasing idling plutonium quantities and
would make the fuel shortage worse overall.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare.eps}
    \caption{
        Separated Pu inventories for all four cases.  Differences in modeling
        choices are not particularly significant until about year 100 when
        fuel shortages begin.  These differences gradually decay away starting
        in year 140 after fuel shortages have ebbed.  Higher aggregate Pu
        inventory  during the shortage is correlated with both a longer shortage and
        a longer post-shortage recovery period.
    }
    \label{fig:puinv-compare}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare-zoom.eps}
    \caption{
        TODO
    }
    \label{fig:puinv-compare-zoom}
\end{figure*}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/sync-cycle.eps}
    \caption[The cycle staggering effect]{
        In this scenario, reactors are deployed annualy with a refueling cycle
        length (including outage) of 18 months.  So all reactors deployed
        every 3rd year refuel together.  All initial 100 LWRs start out with
        their cycles synced as well.
    }
    \label{fig:sync-cycle}
\end{figure}

@@

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases MF and QF.  However, due to the longer
time steps in case QF, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This increases the discrepency between the two
cases. Eventually the shortage in case MF ends slightly before it does in case
QF (shown most clearly in Figure \ref{fig:puinv-compare-zoom}). This feedback effect occurs because the reactors have a
conversion ratio greater than 1.0.

The horizontal striations visible in the monthly flows for case MI in Figure
\ref{fig:puinv} are a characteristic of the synchronization of reactor
refueling caused by the shortage.  A close-up view of these can be seen in
Figure \ref{fig:puinv-MI}. They begin at the end of the fuel shortage window
when many reactors come online together in groups and are resonances that the
inventory level jumps between.  These resonances gradually disappear as the
simulation progresses farther in time beyond the shortage and refueling cycles
become more staggered.

The in-flows shown for cases MF and QF show a periodic pulsing that begins
during the shortage in about year 120 and is shown in greater detail in Figure \ref{fig:puinv-MF}. This pulsing actually has nothing to
do with the shortage and is caused by the retirement of fast reactors.  The
first fast reactors are deployed in year 35 and have an 80 year lifetime.
When they retire, they discharge an entire core's worth of fuel rather than a
single batch.  After a 7-year cooling period, the first discharged full cores
begin to make it through the recycle loop in year 122.  Another interesting
feature visible in all four cases in Figure \ref{fig:puinv} is the minority
fraction of out-flow points that are higher than the bulk majority. These
higher out-flows occur once every 21 months when new reactors are built because
new reactors require a full core of fresh fuel rather than the single batch
needed when just refueling.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-MI.eps}
    \caption[Separated Pu inventory and flow: Case MI Zoom]{
        Separated Pu inventory and flows for case MI.  At locations where the
        outflow dots lie on/near the inventory level, fuel shortages are
        taking place.  This begins around year 105 and continues until about
        year 140 where Pu withdrawals become lower than standing inventory.
        The horizontal bands in out-flows starting around year 140 are caused
        by many previously offline reactors coming online together in groups.
        These fuel inventory level resonances gradually disappear as the
        simulation progresses.
    }
    \label{fig:puinv-MI}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-MF.eps}
    \caption[Separated Pu inventory and flow: Case MF Zoom]{
        Separated Pu inventory and flows for case MF.  At locations where the
        outflow dots meet the inventory level, fuel shortages are taking
        place.  This begins around year 105 and continues until about year 140
        where Pu withdrawals become lower than standing inventory.  The
        periodic pulsing that begins just after year 120 is caused by the
        first fast reactor retirements; the first fast reactors deployed starting in
        year 35 retire annually after their 80 year lifetime and discharge an entire core's
        worth of fuel rather than the usual single batch for refueling.
    }
    \label{fig:puinv-MF}
\end{figure*}

Figure \ref{fig:puinv-compare} shows the same inventory curves from Figure
\ref{fig:puinv} superimposed. Perhaps counter-intuitively, lower separated
plutonium inventory during the shortage generally indicates fewer unfueled
reactors.  Lower inventory levels mean available Pu is being utilized more
efficiently rather than sitting idle.  Cases QI and QF with their longer time
steps show higher inventories indicating a more severe shortage.  Cases QI and
QF take longer to recover from the shortage and fall a bit behind with respect
to building up Pu stocks due to the delayed contributions of more unfueled
reactors to the separated Pu pool.  This can be seen most clearly in Figure
\ref{fig:puinv-compare-zoom}.  Cases MI and QF recover Pu inventory
post-shortage at about the same time and rate - this is consistent with their
approximately equivalent cumulative offline power capacity curves in Figure
\ref{fig:unfueled}.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare.eps}
    \caption{
        Separated Pu inventories for all four cases.  Differences in modeling
        choices are not particularly significant until about year 100 when
        fuel shortages begin.  These differences gradually decay away starting
        in year 140 after fuel shortages have ebbed.
    }
    \label{fig:puinv-compare}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare-zoom.eps}
    \caption{
        Separated Pu inventories for all four cases zoomed in on the fuel
        shortage years.  Higher aggregate Pu inventory  during the shortage is
        correlated with both a longer shortage and a slower post-shortage
        recovery of Pu inventory.
    }
    \label{fig:puinv-compare-zoom}
\end{figure*}

The reactors in all the scenarios request fresh fuel only when they need it.
While this is an unrealistic behavior for real-world facilities, it is closer
to a globally optimum fuel management strategy.  Keeping on-hand fresh fuel
batches would also have the same effect of increasing idling plutonium
quantities and would make the fuel shortage worse overall.
